---
title: "Lineup Experimentation Design"
author: "Emily Robinson"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
library(knitr)
library(tidyverse)
library(readr)
```

### Line-up Task

```{r line-up-examples, message=FALSE, warning=FALSE, fig.width = 12, fig.height = 6, fig.align = "center", echo = F}
library(cowplot)
p1 <- ggdraw() + draw_image("linear-lineup-example.png", scale = 1) + theme_void()
p2 <- ggdraw() + draw_image("log-lineup-example.png", scale = 1) + theme_void()

plot_grid(p1, p2)
```

### Treatment Design

2 test_params x (18 test param_value levels + 6 rorschach param_values levels)

1. Test_param (2 levels)
    - Log Scale
    - Linear Scale
    
2. Param_value (18 test levels + 6 rorschach levels) This is made up of a combination of curvature (Hard / Medium / Easy), variability (Low / High), and panel (Target / Null)

    - Within low / high variability (2 variabilities x 6 combinations of the 3 curvature difficulties = 12 test treatments)
        - Hard Target / Easy Null
        - Hard Target / Medium Null
        - Medium Target / Easy Null
        - Medium Target / Medium Null
        - Easy Target / Medium Null
        - Easy Target / Hard Null
        
    - Within hard / medium / easy curvature (3 curvatures x 2 combinations of the 2 varaibilities = 6 test treatments)
        - Low Target / High Null
        - High Target / Low Null
        
    - For the rorschach treatments (6 levels of curvature x variability where the target and null are the same param_value)
        - Hard Low
        - Hard High
        - Medium Low
        - Medium High
        - Easy Low
        - Easy High

    - param_value labels are given by: target-*Curvature*-*Variability*_null-*Curvature*-*Variability*_*rorschach*
        - Example: target-E-Lv_null-H-Lv_r0 indicates the target panel gets the easy curvature and low variability parameters while the null panels gets the hard curvature and low variability parameters and this is not a rorschach plot (r0).
        - Example: target-E-Lv_null-E-Lv_r1 indicates the lineup is a rorschach (r1) with easy curvature and low varaibility parameters for both the target and null panels. 
    

- 2 reps (datasets) were created for each param_value (total of 2 x (18 + 6) = 48 datasets)
- Each dataset was used in a linear scale lineup and a log scale lineup (total of 2 x 48 = 96 lineups)

### Experiment Design

Incomplete Block Design with a Split Plot

- Each individual recieved a total of 20 plots (18 test plots and 2 rorschach plots)
- Of the 18 test plots
    
    - 9 of the 18 test param_values were randomly selected for each individual (i.e. IBD with block size of 9, 18 treatment levels, and "unlimmited" blocks)
    - Each of the 9 selected param_values x test_param (log/linear) were then randomly assigned a dataset rep. (i.e. each of the 9 param values was given on both the linear/log scale but the dataset (rep) may have differed between the linear/log scale).
    
- Similarly, of the 2 rorschach plots

    - 1 of the 6 rorschach param_values was randomly selected for each individual and each test_param (log/linear scale) was randomly assigned a dataset (rep).
    
Therefore,

- Individual = block
- Param_value = whole plot factor
- Test_param = split plot factor
- Rep (*currently ignoring this, but would this be a row-column blocking type of design?*)

### Preliminary Analysis

For the $i^{th}$ param_value on the $j^{th}$ scale answered by the $k^{th}$ participant, $y_{ijk} = 1$ if the participant is correct and $y_{ijk} = 0$ if the participant is not correct.

Then $y_{ijk} | b_k wp_j ~ Bin(1,p_{ijk})$ where $\eta_{ijk} = \ln\left(\frac{p_{ijk}}{1-p_{ijk}}\right) = \eta + \tau_i + \delta_j + \tau\delta_{ij} + b_k + wp_{ik}$ with $\tau_i$ denoting the effect of the parameter value and $\delta_j$ denoting the effect of the test parameter. 

<br> 

*Assuming n participants per t param_values & not accounting for dataset (rep)*

| Source of Variation                       | DF                    |
|-------------------------------------------|-----------------------|
| participant - *Block Variance*            | n-1                   |
| param_value                               | t-1                   |
| participant x param_value - *WP Variance* | (n-1)(t-1)            |
| test_param                                | (2-1) = 1             |
| param_value x test_param                  | (t-1)(2-1) = (t-1) |
| test_param x participant (param_value) - *overdispersion?*    | (2-1)(n-1)(t)     |

### Preliminary Results

### Raw Values

```{r raw-plot, fig.width = 9, fig.height = 9, echo = F}
raw_data <- read.csv('../../data/graphics-group-09.17.2020.csv')
raw_data %>%
  ggplot(aes(x = test_param, y = correct, color = test_param)) +
  geom_jitter(width = 0.15, height = 0.15, alpha = 0.9) +
  facet_wrap(~param_value) +
  theme_bw() +
  theme(aspect.ratio = 1) +
  scale_color_brewer(palette = "Paired")
```

#### Parameter Value Summary

```{r param_value-summary, fig.width = 9, fig.height = 6, message = F, warning = F}
raw_data %>% 
  group_by(param_value, test_param) %>%
  summarise(plots_correct = sum(correct),
            participant_count = n()) %>%
  mutate(prop_correct = plots_correct/participant_count) %>%
  filter(participant_count > 3) %>%
ggplot(aes(x = test_param, y = prop_correct, fill = test_param)) +
  geom_bar(stat = "identity") +
  facet_wrap(~param_value, ncol = 5) +
  theme_bw() +
  theme(aspect.ratio = 1) +
  scale_fill_brewer(palette = "Paired") +
  scale_y_continuous(limits = c(0,1), breaks = seq(0,1,0.25))
```

#### Participant Summary

```{r participant-summary, fig.height=6, fig.width=9, message=FALSE, warning=FALSE}
raw_data %>% 
  group_by(nick_name, test_param) %>%
  filter(rorschach == 0) %>%
  summarise(plots_correct = sum(correct),
            plot_count = n()) %>%
  mutate(prop_correct = plots_correct/plot_count) %>%
  filter(plot_count > 6) %>%
ggplot(aes(x = test_param, y = prop_correct, fill = test_param)) +
  geom_bar(stat = "identity") +
  facet_wrap(~nick_name, ncol = 4) +
  theme_bw() +
  theme(aspect.ratio = 1) +
  scale_fill_brewer(palette = "Paired") +
  scale_y_continuous(limits = c(0,1), breaks = seq(0,1,0.25))
```