---
title: "Simulation Options (Exponential Multiplicative)"
author: "Emily Robinson"
date: "9/1/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message = FALSE, warning = FALSE, fig.width = 9, fig.height = 6)
library(ggplot2)
library(scales)
library(tidyverse)
library(gridExtra)
```

# Does not include theta

### Heuristic Simulation

$y = \alpha e^{\beta x}$

```{r heuristicSimulation_noTheta, message=FALSE, warning=FALSE}
expSim <- function(xMid, xRange = c(1,20), yRange = c(1,100)){
  
  lineData   <- tibble(xLine = seq(xRange[1],xRange[2],0.1),
                       yLine = ((yRange[1]-yRange[2])/(xRange[2]-xRange[1]))*(xLine-xRange[1])+yRange[2])
  pointsData <- tibble(xPoint = c(xRange[1], xMid, xRange[2]),
                       yPoint = c(yRange[1], lineData$yLine[lineData$xLine == xMid], yRange[2]))
  
  
  lm.fit <- lm(log(yPoint) ~ xPoint, data = pointsData)
  alpha.0  <- exp(coef(lm.fit)[1]) %>% as.numeric()
  beta.0   <- coef(lm.fit)[2] %>% as.numeric()
  start <- list(alpha = alpha.0, beta = beta.0)
  nonlinear.fit   <- nls(yPoint ~ alpha * exp(beta * xPoint), data = pointsData, start = start)
  alpha <- coef(nonlinear.fit)[1] %>% as.numeric()
  beta <- coef(nonlinear.fit)[2] %>% as.numeric()
  
  expData <- tibble(x = seq(xRange[1],xRange[2],0.1),
                    y = alpha*exp(beta*x))
  return(list(expData = expData, lineData = lineData, pointsData = pointsData))
}

expData <- tibble(xMid = c(11, 13, 15)) %>%
  mutate(data = map(xMid, expSim)) %>%
  unnest(data) %>%
  unnest(data) %>%
  mutate(xMid = factor(xMid))

linearPlot1 <- expData %>%
  ggplot(aes(x = x, y = y, group = xMid, color = xMid)) + 
  geom_line() +
  geom_line(aes(x = xLine, y = yLine), color = "gray") +
  geom_point(aes(x = xPoint, y = yPoint), color = "black") +
  theme_bw() +
  scale_color_brewer(palette = "Paired")

logPlot1 <- expData %>%
  ggplot(aes(x = x, y = y, group = xMid, color = xMid)) + 
  geom_line() +
  geom_point(aes(x = xPoint, y = yPoint), color = "black") +
  geom_line(aes(x = xLine, y = yLine), color = "gray") +
  theme_bw() +
  scale_color_brewer(palette = "Paired") +
  scale_y_continuous(trans = "log10",
                     breaks = trans_breaks("log10", function(x) 10^x),
                     labels = trans_format("log10", math_format(10^.x)))

```

$y = \alpha e^{\beta x + \epsilon}$ with $\alpha = \frac{\hat\alpha}{e^{\sigma^2/2}}$

```{r heuristicSimulation_noTheta_error, message=FALSE, warning=FALSE}
expSim <- function(xMid, sigma, N = 20, xRange = c(1,20), yRange = c(1,100)){
  
  lineData   <- tibble(xLine = seq(xRange[1],xRange[2],0.1),
                       yLine = ((yRange[1]-yRange[2])/(xRange[2]-xRange[1]))*(xLine-xRange[1])+yRange[2])
  pointsData <- tibble(xPoint = c(xRange[1], xMid, xRange[2]),
                       yPoint = c(yRange[1], lineData$yLine[lineData$xLine == xMid], yRange[2]))
  
  
  lm.fit <- lm(log(yPoint) ~ xPoint, data = pointsData)
  alpha.0  <- exp(coef(lm.fit)[1]) %>% as.numeric()
  beta.0   <- coef(lm.fit)[2] %>% as.numeric()
  start <- list(alpha = alpha.0, beta = beta.0)
  nonlinear.fit   <- nls(yPoint ~ alpha * exp(beta * xPoint), data = pointsData, start = start)
  alpha <- (coef(nonlinear.fit)[1] %>% as.numeric())/(exp((sigma^2)/2))
  beta <- coef(nonlinear.fit)[2] %>% as.numeric()
  
  expData <- tibble(x = seq(xRange[1],xRange[2], length.out = N),
                    y = alpha*exp(beta*x + rnorm(N,0,sigma)))
  return(list(expData = expData, lineData = lineData, pointsData = pointsData))
}

expData <- tibble(xMid = c(11, 13, 15),
                  sigma = c(0.07, 0.14, 0.25)) %>%
  mutate(data = map2(xMid, sigma, expSim)) %>%
  unnest(data) %>%
  unnest(data) %>%
  mutate(xMid = factor(xMid))

linearPlot1_error <- expData %>%
  ggplot(aes(x = x, y = y, group = xMid, color = xMid)) + 
  geom_point() +
  geom_line(aes(x = xLine, y = yLine), color = "gray") +
  geom_point(aes(x = xPoint, y = yPoint), color = "gray") +
  theme_bw() +
  scale_color_brewer(palette = "Paired")

logPlot1_error <- expData %>%
  ggplot(aes(x = x, y = y, group = xMid, color = xMid)) + 
  geom_point() +
  geom_point(aes(x = xPoint, y = yPoint), color = "gray") +
  geom_line(aes(x = xLine, y = yLine), color = "gray") +
  theme_bw() +
  scale_color_brewer(palette = "Paired") +
  scale_y_continuous(trans = "log10",
                     breaks = trans_breaks("log10", function(x) 10^x),
                     labels = trans_format("log10", math_format(10^.x)))
```

```{r plot1}
grid.arrange(linearPlot1, logPlot1, linearPlot1_error, logPlot1_error, ncol = 2)
```

### Algebraic Simulation

$y = \alpha e^{\beta x}$ with $\alpha = \frac{yMax}{e^{\beta xMax}}$

```{r originalSim_max, message=FALSE, warning=FALSE}
expSim <- function(beta, xRange = c(1,20), yRange = c(1,100)){
  alpha   <- (yRange[2]/exp(beta*xRange[2]))
  expData <- tibble(x = seq(xRange[1],xRange[2],1),
                    y = alpha*exp(beta*x))
  return(expData)
}

expData <- tibble(beta = c(0.07, 0.15, 0.3)) %>%
           mutate(data = map(beta, expSim)) %>%
           unnest(data) %>%
           mutate(beta = factor(beta))

linearPlot2 <- expData %>%
  ggplot(aes(x = x, y = y, group = beta, color = beta)) + 
  geom_line() +
  theme_bw() +
  scale_color_brewer(palette = "Paired")

logPlot2 <- expData %>%
  ggplot(aes(x = x, y = y, group = beta, color = beta)) + 
  geom_line() +
  theme_bw() +
  scale_color_brewer(palette = "Paired") +
  scale_y_continuous(trans = "log10",
                     breaks = trans_breaks("log10", function(x) 10^x),
                     labels = trans_format("log10", math_format(10^.x)))
```


$y = \alpha e^{\beta x + \epsilon}$ with $\alpha = \frac{yMax}{e^{\beta xMax+\sigma^2/2}}$

```{r originalSim_max_error, message=FALSE, warning=FALSE}
expSim <- function(beta, sigma, N = 30, xRange = c(1,20), yRange = c(1,100)){
  alpha <- (yRange[2]/exp(beta*xRange[2]))/(exp((sigma^2)/2))
  expData <- tibble(x = seq(xRange[1],xRange[2], length.out = N),
                    y = alpha*exp(beta*x + rnorm(N,0,sigma)))
  return(expData)
}

expData <- tibble(beta = c(0.3, 0.15, 0.07),
                  sigma = c(0.3, 0.12, 0.05)) %>%
  mutate(data = map2(beta, sigma, expSim)) %>%
  unnest(data) %>%
  mutate(beta = factor(beta))

linearPlot2_error <- expData %>%
  ggplot(aes(x = x, y = y, group = beta, color = beta)) + 
  geom_point() +
  theme_bw() +
  scale_color_brewer(palette = "Paired")

logPlot2_error <- expData %>%
  ggplot(aes(x = x, y = y, group = beta, color = beta)) + 
  geom_point() +
  theme_bw() +
  scale_color_brewer(palette = "Paired") +
  scale_y_continuous(trans = "log10",
                     breaks = trans_breaks("log10", function(x) 10^x),
                     labels = trans_format("log10", math_format(10^.x)))
```

```{r plot2}
grid.arrange(linearPlot2, logPlot2, linearPlot2_error, logPlot2_error, ncol = 2)
```

# Includes theta

### Heuristic Simulation

$y = \alpha e^{\beta x} + \theta$

```{r heuristicSimulation_yesTheta, message=FALSE, warning=FALSE}
expSim <- function(xMid, xRange = c(0,20), yRange = c(1,100)){
  
  lineData   <- tibble(xLine = seq(xRange[1],xRange[2],0.1),
                       yLine = ((yRange[1]-yRange[2])/(xRange[2]-xRange[1]))*(xLine-xRange[1])+yRange[2])
  pointsData <- tibble(xPoint = c(xRange[1], 
                                  (xMid-0.1), 
                                  (xMid+0.1), 
                                  xRange[2]
                                  ),
                       yPoint = c(yRange[1], 
                                  lineData$yLine[lineData$xLine == xMid], 
                                  lineData$yLine[lineData$xLine == xMid], 
                                  yRange[2]
                                  )
                       )
  
  
  lm.fit <- lm(log(yPoint) ~ xPoint, data = pointsData)
  alpha.0  <- exp(coef(lm.fit)[1]) %>% as.numeric()
  beta.0   <- coef(lm.fit)[2] %>% as.numeric()
  theta.0 <- min(pointsData$yPoint) * 0.5  
  start <- list(alpha = alpha.0, beta = beta.0, theta = theta.0)
  nonlinear.fit   <- nls(yPoint ~ alpha * exp(beta * xPoint) + theta , data = pointsData, start = start)
  alpha <- coef(nonlinear.fit)[1] %>% as.numeric()
  beta <- coef(nonlinear.fit)[2] %>% as.numeric()
  theta <- coef(nonlinear.fit)[3] %>% as.numeric()

  expData <- tibble(x = seq(xRange[1],xRange[2],0.1),
                    y = alpha*exp(beta*x) + theta)
  return(list(expData = expData, lineData = lineData, pointsData = pointsData))
}

expData <- tibble(xMid = c(11.8, 13, 14.5)) %>%
  mutate(data = map(xMid, expSim)) %>%
  unnest(data) %>%
  unnest(data) %>%
  mutate(xMid = factor(xMid))

linearPlot3 <- expData %>%
  ggplot(aes(x = x, y = y, group = xMid, color = xMid)) + 
  geom_line() +
  geom_point(aes(x = xPoint, y = yPoint), color = "black") +
  geom_line(aes(x = xLine, y = yLine), color = "gray") +
  theme_bw() +
  scale_color_brewer(palette = "Paired")

logPlot3 <- expData %>%
  ggplot(aes(x = x, y = y, group = xMid, color = xMid)) + 
  geom_line() +
  geom_point(aes(x = xPoint, y = yPoint), color = "black") +
  geom_line(aes(x = xLine, y = yLine), color = "gray") +
  theme_bw() +
  scale_color_brewer(palette = "Paired") +
  scale_y_continuous(trans = "log10",
                     breaks = trans_breaks("log10", function(x) 10^x),
                     labels = trans_format("log10", math_format(10^.x)))

```

$y = \alpha e^{\beta x + \epsilon} + \theta$ with $\alpha = \frac{\hat\alpha}{e^{\sigma^2/2}}$

```{r heuristicSimulation_yesTheta_error, message=FALSE, warning=FALSE}
expSim <- function(xMid, sigma, N = 20, xRange = c(0,20), yRange = c(1,100)){
  
  lineData   <- tibble(xLine = seq(xRange[1],xRange[2],0.1),
                       yLine = ((yRange[1]-yRange[2])/(xRange[2]-xRange[1]))*(xLine-xRange[1])+yRange[2])
  pointsData <- tibble(xPoint = c(xRange[1], (xMid-0.1), (xMid+0.1), xRange[2]),
                       yPoint = c(yRange[1], lineData$yLine[lineData$xLine == xMid], lineData$yLine[lineData$xLine == xMid], yRange[2]))
  
  
  lm.fit <- lm(log(yPoint) ~ xPoint, data = pointsData)
  alpha.0  <- exp(coef(lm.fit)[1]) %>% as.numeric()
  beta.0   <- coef(lm.fit)[2] %>% as.numeric()
  theta.0 <- min(pointsData$yPoint) * 0.5  
  start <- list(alpha = alpha.0, beta = beta.0, theta = theta.0)
  nonlinear.fit   <- nls(yPoint ~ alpha * exp(beta * xPoint) + theta , data = pointsData, start = start)
  alpha <- (coef(nonlinear.fit)[1] %>% as.numeric())/(exp((sigma^2)/2))
  beta <- coef(nonlinear.fit)[2] %>% as.numeric()
  theta <- coef(nonlinear.fit)[3] %>% as.numeric()

  expData <- tibble(x = seq(xRange[1],xRange[2], length.out = N),
                    y = alpha*exp(beta*x + rnorm(N,0,sigma)) + theta)
  return(list(expData = expData, lineData = lineData, pointsData = pointsData))
}

expData <- tibble(xMid = c(11.8, 13, 14.5),
                  sigma = c(0.07, 0.14, 0.25)) %>%
  mutate(data = map2(xMid, sigma, expSim)) %>%
  unnest(data) %>%
  unnest(data) %>%
  mutate(xMid = factor(xMid))

linearPlot3_error <- expData %>%
  ggplot(aes(x = x, y = y, group = xMid, color = xMid)) + 
  geom_point() +
  geom_point(aes(x = xPoint, y = yPoint), color = "gray") +
  geom_line(aes(x = xLine, y = yLine), color = "gray") +
  theme_bw() +
  scale_color_brewer(palette = "Paired")

logPlot3_error <- expData %>%
  ggplot(aes(x = x, y = y, group = xMid, color = xMid)) + 
  geom_point() +
  geom_point(aes(x = xPoint, y = yPoint), color = "gray") +
  geom_line(aes(x = xLine, y = yLine), color = "gray") +
  theme_bw() +
  scale_color_brewer(palette = "Paired") +
  scale_y_continuous(trans = "log10",
                     breaks = trans_breaks("log10", function(x) 10^x),
                     labels = trans_format("log10", math_format(10^.x)))
```

```{r plot3}
grid.arrange(linearPlot3, logPlot3, linearPlot3_error, logPlot3_error, ncol = 2)
```

### Algebraic Simulation

$y = \alpha e^{\beta x} + \theta$ with $\theta = \frac{yMax - yMin\times e^{\beta (xMax - xMin)}}{1 - e^{\beta xMax}}$ and $\alpha=\frac{yMin - \theta}{e^{\beta xMin}}$

```{r originalSim_min-max}
expSim <- function(beta, xRange = c(0,20), yRange = c(1,100)){
  theta <- (yRange[2]-yRange[1]*(exp(beta*(xRange[2]-xRange[1]))))/(1-exp(beta*xRange[2]))
  alpha <- (yRange[1]-theta)/(exp(beta*xRange[1]))
  expData <- tibble(x = seq(xRange[1],xRange[2],1),
                    y = alpha*exp(beta*x)+theta)
  return(expData)
}

expData <- tibble(beta = c(0.07, 0.15, 0.3)) %>%
  mutate(data = map(beta, expSim)) %>%
  unnest(data) %>%
  mutate(beta = factor(beta))

linearPlot4 <- expData %>%
  ggplot(aes(x = x, y = y, group = beta, color = beta)) + 
  geom_line() +
  theme_bw() +
  scale_color_brewer(palette = "Paired")

logPlot4 <- expData %>%
  ggplot(aes(x = x, y = y, group = beta, color = beta)) + 
  geom_line() +
  theme_bw() +
  scale_color_brewer(palette = "Paired") +
  scale_y_continuous(trans = "log10",
                     breaks = trans_breaks("log10", function(x) 10^x),
                     labels = trans_format("log10", math_format(10^.x)))
```

$y = \alpha e^{\beta x + \epsilon} + \theta$ with $\theta = \frac{yMax - yMin\times e^{\beta (xMax - xMin)}}{1 - e^{\beta xMax}}$ and $\alpha=\frac{yMin - \theta}{e^{\beta xMin+\sigma^2/2}}$

```{r originalSim_min-max_error}
expSim <- function(beta, sigma, N = 30, xRange = c(0,20), yRange = c(1,100)){
  theta <- (yRange[2]-yRange[1]*(exp(beta*(xRange[2]-xRange[1]))))/(1-exp(beta*xRange[2]+sigma^2/2))
  alpha <- (yRange[1]-theta)/(exp(beta*xRange[1]+sigma^2/2))
  expData <- tibble(x = seq(xRange[1],xRange[2], length.out = N),
                    y = alpha*exp(beta*x + rnorm(N,0,sigma)) + theta)
  return(expData)
}

expData <- tibble(difficulty = c("Easy", "Medium", "Hard"),
                  beta = c(0.3, 0.15, 0.07),
                  sigma = c(0.3, 0.12, 0.05)) %>%
  mutate(data = map2(beta, sigma, expSim)) %>%
  unnest(data) %>%
  mutate(beta  = factor(beta),
         sigma = factor(sigma))

linearPlot4_error <- expData %>%
  ggplot(aes(x = x, y = y, group = beta, color = beta)) + 
  geom_point() +
  theme_bw() +
  scale_color_brewer(palette = "Paired")

logPlot4_error <- expData %>%
  ggplot(aes(x = x, y = y, group = beta, color = beta)) + 
  geom_point() +
  theme_bw() +
  scale_color_brewer(palette = "Paired") +
  scale_y_continuous(trans = "log10",
                     breaks = trans_breaks("log10", function(x) 10^x),
                     labels = trans_format("log10", math_format(10^.x)))
```

```{r plot4}
grid.arrange(linearPlot4, logPlot4, linearPlot4_error, logPlot4_error, ncol = 2)
```